TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages

### 2021-03-26 article.tex
paper TTT: detalles en la parte anterior a ilustraciones

\en{As an instance, let's consider a winning case ($p_i > p_j$) using a Gaussian prior (i.e. $\N(\,s\,|\,\mu, \sigma^2)$) for each of the skills.}

@@

\en{The normalized posterior can be found dividing the right hand by the evidence $P(r)$.}

@@

\es{Es interesante notar las similitudes y diferencias entre el likelihood y el evidencia.}

@@

    \en{The proportinal posterior is obtained as the product of the prior (Gaussian) and the likelihood (cumulative Gaussian).}

@@

    \en{The distributions are not neccesary on the same scale: the prior integrates 1, while the likelihood goes from 0 to 1.}

@@

\en{At the region of very low skill hypotheses, where the winning result would have generated huge surprise ($\lim_{s_i \to -\infty}P(r|s_i) = 0$), the posterior receive no density from the prior.}

@@

\en{It is important to stress that the posterior, although similar, is not a Gaussian distribution, preventing us from using equation~\ref{eq:posterior_win} iteratively.}
\es{Es importante remarcar que la posterior, aunque se parezca, no es una distribuci\'on gaussiana, lo que nos impedir\'a usar la ecuaci\'on~\ref{eq:posterior_win} iterativamente.}

@@

\en{The success of the TrueSkill solution~\citep{Herbrich2007} is based on the application of an efficient method for computing the Gaussian distribution that guarantees the approximation of the posterior (section~\ref{sec:approximate_posterior}),}

@@

\en{This method allow as to efficiently apply equation~\eqref{eq:posterior_win} iteratively over a sequence of observations, which would otherwise be infeasible.}

@@

\en{If we consider the likelihood as a filter of the prior, each posterior can be seen as an accumulation of all previous filters.}

@@

\en{The class \texttt{History} is used to compute the posteriors and evidence of a sequence of events.}
\es{La clase \texttt{History} se usa para computar los posteriors y la evidencia de secuencias de eventos.}
%
\en{In the first example, we instantiate the class \texttt{History} with three players (\texttt{\small "a", "b", "c"}) and three games.}
\es{En el primer ejemplo, inicializamos una la clase \texttt{History} con tres jugadores (\texttt{"a","b","c"}) y tres partidas.}
%
\en{In the first game \texttt{"a"} beats \texttt{"b"}, in the second game \texttt{"b"} beats \texttt{"c"}, and in the third game \texttt{"c"} beats \texttt{"a"}.}
\es{En la primera partida \texttt{"a"} le gana a \texttt{"b"}, en la segunda \texttt{"b"} le gana a \texttt{"c"} y en la tercera \texttt{"c"} le gana a \texttt{"a"}.}
%
\en{In brief, all agents win one game and lose the other.}
\es{En resumen, todos los agentes ganan una partida y pierden la otra.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:history, caption={\en{Initialization of a \texttt{History}'s instance with a three events sequence.}\es{Inicialización de una instancia de \texttt{History} con una secuencia de tres eventos}}, belowskip=-1.0 \baselineskip, aboveskip=0.1cm]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia},belowskip=-0.77 \baselineskip]
c1 = [["a"],["b"]]
c2 = [["b"],["c"]]
c3 = [["c"],["a"]]
composition = [c1, c2, c3]
\end{lstlisting}
  \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python},belowskip=-0.77 \baselineskip]
c1 = [["a"],["b"]]
c2 = [["b"],["c"]]
c3 = [["c"],["a"]]
composition = [c1, c2, c3]
\end{lstlisting}
   \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r},belowskip=-0.77 \baselineskip]
c1 = list(c("a"),c("b"))
c2 = list(c("b"),c("c"))
c3 = list(c("c"),c("a"))
composition = list(c1,c2,c3)
\end{lstlisting}
\end{paracol}
\begin{lstlisting}[backgroundcolor=\color{all}]
h = History(composition, gamma=0.0)
\end{lstlisting}
%
\en{where the variables \texttt{c1}, \texttt{c2}, and \texttt{c3} model the composition of each game using the names of the agents (i.e. identifiers), the variable \texttt{composition} is a list containing the three events, and the parameter \texttt{gamma} specifies how much the skills change over time (where 0 means no change).}
\es{donde la variable \texttt{c1}, \texttt{c2} y \texttt{c3} modela la composici\'on de cada partida usando los nombres de los agentes (i.e. identificadores), la variable \texttt{composition} es una lista que contiene las tres partidas, y el par\'ametro \texttt{gamma} especifica cuánto cambian las habilidad en el tiempo (donde 0 significa ningún cambio).}
%
\en{The results are implicitly defined by the order in which the game compositions are initialized: the teams appearing first in the list beat those appearing later.}
\es{El resultado queda definido implícitamente por el orden en el que la composiciones de las partidas fueron inicializadas: los equipos que aparecen primero en la lista vencen a los que aparecen después.}
%
\en{The rest of the parameters are initialized with the default values seen in code~\ref{lst:parameters}.}
\es{El resto de los parámetros se inicializa con los valores por defecto vistos en el código~\ref{lst:parameters}.}

@@

\en{TrueSkill Through Time not only returns correct estimates (the same for all players), they also have less uncertainty.}

@@

\en{The class \texttt{History} receives this four parameters and initializes the target player using the default values and a dynamic uncertainty \texttt{gamma=0.015}.}
\es{La clase \texttt{History} recibe estos cuatro parámetros e inicializa al agente objetivo con los valores por defecto y un incertidumbre dinámico \texttt{gamma=0.015}.}
%
\en{Using the method \texttt{convergence()} we obtain the TrueSkill Through Time estimates and the target's learning curve.}
\es{Usando el m\'etodo \texttt{convergence()} obtenemos las estimaciones de TrueSkill Through Time y la curva de aprendizaje del agente objetivo.}
%
\en{Because the estimates depend on random results, we decided to repeat procedure~\ref{lst:estimating_the_simulated} to account for their variability.}
\es{Debido a que las estimaciones dependen de los resultados aleatorios, para tener en cuenta su variabilidad decidimos repetir el procedimiento~\ref{lst:estimating_the_simulated}.}
%
\en{The Fig~\ref{fig:logistic} shows the evolution of the true (solid line) and estimated (dotted line) learning curves of the target player.}
\es{La figura~\ref{fig:logistic} muestra la evolución de la curva de aprendizaje real (línea sólida) y estimada (línea punteada) del jugador objetivo.}
%
\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
    \en{\includegraphics[page={3},width=1.\linewidth]{figures/logistic}}
    \es{\includegraphics[page={4},width=1.\linewidth]{figures/logistic}}
    \caption{
    \en{Mean estimates}
    \es{Varias estimaciones medias}
    }
    \label{fig:logistic_mu}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
    \en{\includegraphics[page={1},width=1.\linewidth]{figures/logistic}}
    \es{\includegraphics[page={2},width=1.\linewidth]{figures/logistic}}
    \caption{
    \en{Uncertainty of an estimate}
    \es{Incertidumbre de una estimación}
    }
    \label{fig:logistic_sigma}
    \end{subfigure}
    \caption{
    \en{True and estimated learning curves of a new player that joins a large community of already known players.}
    \es{Curva de aprendizaje sintética y estimada de un jugador nuevo que se incorpora a una comunidad grande de jugadores ya conocidos.}
    %
    \en{The solid line represents the true skill of the target player while the dashed lines show the mean estimates.}
    \es{La línea continua representa la habilidad real del jugador objetivo, mientras que la línea discontinua muestra la estimación media.}
    %
    \en{The dark and light gray areas show onnce and twice the uncertainty of one of the estimates.}
    \es{Las zonas oscuras y claras ilustra una y dos veces la incertidumbre de la estimación.}
    }
    \label{fig:logistic}
\end{figure}
%
\en{The estimated learning curves remain close to the actual skill during the whole evolution as can be seen in Fig~\ref{fig:logistic_mu}.}
\es{En la figura~\ref{fig:logistic_mu} vemos que las medias de las curvas de aprendizaje estimada se mantienen siempre cerca de la habilidad real durante toda la evolución.}
%
\en{For the Fig~\ref{fig:logistic_sigma}, we select one of the estimated learning curves and present its uncertainty, showing that the actual learning curve is contained in the uncertainty interval.}
\es{En la figura~\ref{fig:logistic_sigma} vemos que la curva real esta siempre contenida en el intervalo de incertidumbre de cualquiera de las estimaciones.}
%
\en{This example exhibit that TrueSkill Through Time correctly estimates and follows the skill evolution of a new player.}
\es{Este ejemplo evidencia que TrueSkill Through Time estima y sigue correctamente la evolución de las habilidades de los jugador nuevos.}

@@

\en{In this last example we use the entire history of matches of the Association of Tennis Professionals (ATP).}
\es{En este último ejemplo utilizamos toda la historia de partidas de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players.}
\es{La base de datos cuenta con \num{447000} mil partidas que van desde el año 1915 hasta el 2020, en las que participan \num{19000} jugadores.}
%
\en{The information contained in the \texttt{csv} files has both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{La información contenida en los archivos \texttt{csv} contiene tanto partidas simples como dobles, indicado por la columna \texttt{double}.}
%
\en{Each game has a unique identifier (i.e. \texttt{match\_id}) and tournament's round number (i.e. \texttt{round\_number}) where 0 represents the final game, 1 the semi-final, and so on.}
\es{Tenemos un identificador por la partida (\texttt{match\_id}) y el número de ronda (\texttt{round\_number}) donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains four columns for the players' identifiers in the game (the last two players are null in the case of a single): for example \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{Contiene cuatro columnas correspondiente a los identificadores de los jugadores y cuatro columnas correspondiente a sus nombres, por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador (los segundos jugadores tienen valores nulos en partidas simples) y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because it is more efficient than \proglang{Python} and \proglang{R} versions, which can be also found in the appendix \ref{AGREGA-CODE}.}
\es{Aquí mostramos solo la versión de \proglang{Julia} debido a que es la más eficiente, las versiones de \proglang{Python} y \proglang{R} se pueden entonrar en el anexo \ref{AGREGA-CODIGO}.}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")

@@

\begin{lstlisting}[backgroundcolor=\color{white},label=lst:performance, caption={\en{Computing the indivudal prior performance}\es{Computando el desempeño individual a priori}}, belowskip=-1.0 \baselineskip, aboveskip=0.0 \baselineskip]

@@

\subsection{\en{Elementary draw model}\es{Modelo b\'asico de empates}} \label{sec:empate}

>>>>

\en{As an example, we consider a winning case ($p_i > p_j$) using a Gaussian prior (i.e. $\N(\,s\,|\,\mu, \sigma^2)$) for each of the skills.}

@@

\en{The normalized posterior is found by dividing the right hand by the evidence $P(r)$.}

@@

\es{Es interesante notar las similitudes y diferencias entre el likelihood y la evidencia.}

@@

    \en{The proportional posterior is obtained as the product of the prior (Gaussian) and the likelihood (cumulative Gaussian).}

@@

    \en{The distributions are not necessarily on the same scale: the prior integrates to $1$, while the likelihood goes from $0$ to $1$.}

@@

\en{At the region of very low skill hypotheses, a win would generate a great surprise ($\lim_{s_i \to -\infty}P(r|s_i) = 0$), and the posterior receives no density from the prior.}

@@

\en{It is important to stress that the posterior, although similar, is not a Gaussian distribution, preventing us from using equation~\eqref{eq:posterior_win} iteratively.}
\es{Es importante remarcar que la posterior, aunque se parezca, no es una distribuci\'on gaussiana, lo que nos impedir\'a usar la ecuaci\'on~\eqref{eq:posterior_win} iterativamente.}

@@

\en{The success of the TrueSkill solution~\citep{Herbrich2007} is based on the usage of an efficient method for computing the Gaussian distribution that guarantees the approximation of the posterior~(see section~\ref{sec:approximate_posterior}),}

@@

\en{This method allows us to efficiently apply equation~\eqref{eq:posterior_win} iteratively over a sequence of observations, which would otherwise be infeasible.}

@@

\en{If we consider the likelihood as a filter of the prior, each posterior is the accumulation of all previous filters.}

@@

\en{The class \texttt{History} is used to compute the posteriors and evidence of a sequence of events.}
\es{La clase \texttt{History} se usa para computar los posteriors y la evidencia de secuencias de eventos.}
%
\en{In the first example, we instantiate the class \texttt{History} with three players (\texttt{\small "a", "b", "c"}) and three games.}
\es{En el primer ejemplo, inicializamos una la clase \texttt{History} con tres jugadores (\texttt{"a","b","c"}) y tres partidas.}
%
\en{In the first game \texttt{"a"} beats \texttt{"b"}, in the second game \texttt{"b"} beats \texttt{"c"}, and in the third game \texttt{"c"} beats \texttt{"a"}.}
\es{En la primera partida \texttt{"a"} le gana a \texttt{"b"}, en la segunda \texttt{"b"} le gana a \texttt{"c"} y en la tercera \texttt{"c"} le gana a \texttt{"a"}.}
%
\en{In brief, all agents win one game and lose the other.}
\es{En resumen, todos los agentes ganan una partida y pierden la otra.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:history, caption={\en{Initialization of a \texttt{History}'s instance with a three events sequence}\es{Inicialización de una instancia de \texttt{History} con una secuencia de tres eventos}}, belowskip=-1.0 \baselineskip, aboveskip=0.1cm]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia},belowskip=-0.77 \baselineskip]
c1 = [["a"],["b"]]
c2 = [["b"],["c"]]
c3 = [["c"],["a"]]
composition = [c1, c2, c3]
\end{lstlisting}
  \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python},belowskip=-0.77 \baselineskip]
c1 = [["a"],["b"]]
c2 = [["b"],["c"]]
c3 = [["c"],["a"]]
composition = [c1, c2, c3]
\end{lstlisting}
   \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r},belowskip=-0.77 \baselineskip]
c1 = list(c("a"),c("b"))
c2 = list(c("b"),c("c"))
c3 = list(c("c"),c("a"))
composition = list(c1,c2,c3)
\end{lstlisting}
\end{paracol}
\begin{lstlisting}[backgroundcolor=\color{all}]
h = History(composition, gamma=0.0)
\end{lstlisting}
%
\en{where the variables \texttt{c1}, \texttt{c2}, and \texttt{c3} model the composition of each game using the names of the agents (i.e. identifiers), the variable \texttt{composition} is a list containing the three events, and the parameter \texttt{gamma} specifies how much the skills change over time (where 0 means no change).}
\es{donde la variable \texttt{c1}, \texttt{c2} y \texttt{c3} modela la composici\'on de cada partida usando los nombres de los agentes (i.e. identificadores), la variable \texttt{composition} es una lista que contiene las tres partidas, y el par\'ametro \texttt{gamma} especifica cuánto cambian las habilidad en el tiempo (donde 0 significa ningún cambio).}
%
\en{The results are implicitly defined by the order in which the game compositions are initialized: the teams appearing first in the list beat those appearing later.}
\es{El resultado queda definido implícitamente por el orden en el que la composiciones de las partidas fueron inicializadas: los equipos que aparecen primero en la lista vencen a los que aparecen después.}
%
\en{The rest of the parameters are initialized with the default values seen in code~\ref{lst:parameters}.}
\es{El resto de los parámetros se inicializa con los valores por defecto vistos en el código~\ref{lst:parameters}.}

@@

\en{TrueSkill Through Time not only returns correct estimates (same for all players), they also have less uncertainty.}

@@

\en{The class \texttt{History} receives this four parameters and initializes the target player using the default values and a dynamic uncertainty \texttt{gamma=0.015}.}
\es{La clase \texttt{History} recibe estos cuatro parámetros e inicializa al agente objetivo con los valores por defecto y un incertidumbre dinámico \texttt{gamma=0.015}.}
%
\en{Using the method \texttt{convergence()} we obtain the TrueSkill Through Time estimates and the target's learning curve.}
\es{Usando el m\'etodo \texttt{convergence()} obtenemos las estimaciones de TrueSkill Through Time y la curva de aprendizaje del agente objetivo.}
%
\en{Because the estimates depend on random results, we decided to repeat procedure~\ref{lst:estimating_the_simulated} to account for their variability.}
\es{Debido a que las estimaciones dependen de los resultados aleatorios, para tener en cuenta su variabilidad decidimos repetir el procedimiento~\ref{lst:estimating_the_simulated}.}
%
\en{The Fig~\ref{fig:logistic} shows the evolution of the true (solid line) and estimated (dotted line) learning curves of the target player.}
\es{La figura~\ref{fig:logistic} muestra la evolución de la curva de aprendizaje real (línea sólida) y estimada (línea punteada) del jugador objetivo.}
%
\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
    \en{\includegraphics[page={3},width=1.\linewidth]{figures/logistic}}
    \es{\includegraphics[page={4},width=1.\linewidth]{figures/logistic}}
    \caption{
    \en{Mean estimates}
    \es{Varias estimaciones medias}
    }
    \label{fig:logistic_mu}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
    \en{\includegraphics[page={1},width=1.\linewidth]{figures/logistic}}
    \es{\includegraphics[page={2},width=1.\linewidth]{figures/logistic}}
    \caption{
    \en{Uncertainty of an estimate}
    \es{Incertidumbre de una estimación}
    }
    \label{fig:logistic_sigma}
    \end{subfigure}
    \caption{
    \en{True and estimated learning curves of a new player that joins a large community of already known players.}
    \es{Curva de aprendizaje sintética y estimada de un jugador nuevo que se incorpora a una comunidad grande de jugadores ya conocidos.}
    %
    \en{The solid line represents the true skill of the target player while the dashed lines show the mean estimates.}
    \es{La línea continua representa la habilidad real del jugador objetivo, mientras que la línea discontinua muestra la estimación media.}
    %
    \en{The dark and light gray areas show one and two times the uncertainty of one of the estimates.}
    \es{Las zonas oscuras y claras ilustra una y dos veces la incertidumbre de la estimación.}
    }
    \label{fig:logistic}
\end{figure}
%
\en{The estimated learning curves remain close to the actual skill during the whole evolution as can be seen in Fig~\ref{fig:logistic_mu}.}
\es{En la figura~\ref{fig:logistic_mu} vemos que las medias de las curvas de aprendizaje estimada se mantienen siempre cerca de la habilidad real durante toda la evolución.}
%
\en{For the Fig~\ref{fig:logistic_sigma}, we select one of the estimated learning curves and present its uncertainty, showing that the actual learning curve is contained in the uncertainty interval.}
\es{En la figura~\ref{fig:logistic_sigma} vemos que la curva real esta siempre contenida en el intervalo de incertidumbre de cualquiera de las estimaciones.}
%
\en{This example exhibit that TrueSkill Through Time correctly estimates and follows the skill evolution of a new player.}
\es{Este ejemplo evidencia que TrueSkill Through Time estima y sigue correctamente la evolución de las habilidades de los jugador nuevos.}

@@

\en{In this last example we use the entire history of matches of the Association of Tennis Professionals (ATP).}
\es{En este último ejemplo utilizamos toda la historia de partidas de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players and is publicly available\footnote{A zip file containing a csv file per year can be downloaded from \url{https://github.com/JeffSackmann/tennis_atp}}.}
\es{La base de datos cuenta con \num{447000} mil partidas que van desde el año 1915 hasta el 2020, en las que participan \num{19000} jugadores.}
%
\en{The information is formatted and stored in a single \texttt{csv} file\footnote{Available at \url{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}} including both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{La información contenida en los archivos \texttt{csv} contiene tanto partidas simples como dobles, indicado por la columna \texttt{double}.}
%
\en{Each game has a unique identifier (i.e. \texttt{match\_id}) and tournament's round number (i.e. \texttt{round\_number}) where 0 represents the final game, 1 the semi-final, and so on.}
\es{Tenemos un identificador por la partida (\texttt{match\_id}) y el número de ronda (\texttt{round\_number}) donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains four columns for the players' identifiers in the game (the last two players are null in the case of a single): for example \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{Contiene cuatro columnas correspondiente a los identificadores de los jugadores y cuatro columnas correspondiente a sus nombres, por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador (los segundos jugadores tienen valores nulos en partidas simples) y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because it is more efficient than \proglang{Python} and \proglang{R} versions, which can be also found in the appendix \ref{AGREGA-CODE}.}
\es{Aquí mostramos solo la versión de \proglang{Julia} debido a que es la más eficiente, las versiones de \proglang{Python} y \proglang{R} se pueden entonrar en el anexo \ref{AGREGA-CODIGO}.}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")

@@

\begin{lstlisting}[backgroundcolor=\color{white},label=lst:performance, caption={\en{Computing the individual prior performance}\es{Computando el desempeño individual a priori}}, belowskip=-1.0 \baselineskip, aboveskip=0.0 \baselineskip]

@@

\subsection{\en{A Basic Draw Model}\es{Modelo b\'asico de empates}} \label{sec:empate}
